I owe a huge debt to my advisor, Nuh Gedik, for all of his support throughout my Ph.D.
Not only is Nuh a brilliant scientist, he is also extremely kind to his students, demonstrating a level of compassion far beyond, I think, what is typical for Ph.D. advisors in Physics.
None of the research in this thesis would have been possible without that.

Nor would it have been possible without the guidance of my undergraduate advisor, Jay Kikkawa, as well as Sebastian Hurtado Parra and Mehmet Noyan, who were graduate students in his group when I was there.
These were the first people in my career who decided to believe in me (based on what, I don't know), and for that I owe them a huge debt.
They are also all excellent experimentalists; almost everything I know about doing research came from them.

Many thanks also to Joe Checkelsky and Liang Fu, who stepped in to serve on my committee and who I very much enjoyed chatting with during my Ph.D.
I am also grateful for the support of my academic advisor, Pablo Jarillo-Herrero.

Anshul Kogar was a postdoc in the Gedik group when I joined in \num{2017}, and took me under his wing when as I was building the initial experimental setup.
Anshul sat through many days, weeks, and months of debugging the setup with me, and taught me how to maintain a positive disposition despite the many challenges we faced during this process.
He was also very kind to me when I was facing some health issues during my first and second years, which I am very grateful for.

Baiqing Lv joined the Gedik group a few years later, and was my officemate and mentor from then until he left in \num{2022}.
Baiqing and I worked jointly on almost all of the projects in this thesis, and he and I spent many hours discussing different interpretations for our results before we landed on the ones presented here.
He was notorious in the Gedik group for being, somehow, always in the know---if you needed a sample, he knew someone who could grow it; if you wanted gossip, academic or not, he somehow had it.
I am very glad to have met Baiqing and am grateful for our continued correspondence, even after he started his job in Shanghai.

Karna Morey and Zongqi Shen, two very perspicacious young people, joined the Gedik group when I was a senior Ph.D. student and helped me a lot with all of the projects presented in this thesis.
Without Karna, we would not have had the ability to automate the polarizers in the experimental setup, which was indisputably necessary for the \ce{CaMn2Bi2} and \ce{CuBr2} projects.
He and Zongqi were also responsible for taking much of the data for these projects when I was travelling or otherwise away from the lab.
After Karna graduated and moved to Stanford for his Ph.D., Zongqi joined the lab has been working with me since then.
In that time, he has not only learned all of the ins and outs of the experimental setup, but also launched a number of new projects which I am excited to see pan out.
I am grateful to both of them not only for their help completing this research, but also for the opportunity to be their mentors, and for their patience as I learned how to do that.

I am grateful also to the rest of the Gedik group for all of their support during my Ph.D; it was truly a pleasure to come to an office every day that was filled with such amazing people, and to be able to enjoy their company throughout the day.
At many times they showed me kindness that they didn't need to---from helping me with experiments to taking me to the hospital when I broke my foot---which had a much greater impact than I think they know.

Thanks also to Monica Wolf, Gerry Miller, and Lars LLorente, who were / are essential in keeping CMX functional at MIT.

To Mom, Dad, Alison, and Dylan, thank you for instilling in me a love for science, supporting, my dreams, and for encouraging me to be the best version of myself that I can be.
Thanks also to Marc, Lisa, Bernie, Antonella, Julius, and Ari for all of your support and for welcoming me into your family.

No amount of thanks is adequte (I'll try anyway) for all of my friends, especially Will, Cliff, Kalina, Hope, Abe, Jeff, and Rachel.
I have no right to have so many amazing people in my life, but I do, and I'm very grateful for you.

Thank you to all of the workers who make life in Boston possible, especially the employees at Chipotle, Life Alive, and all the other killer lunch spots near the office, the custodial and facilities staffs at MIT, the employees at MIT Medical, and the employees / delivery drivers at the various restaurants around town that I would get dinner from.

Finally, to Cami: thank you for coming with me where I wanted to go, and for carrying me when I couldn't go any further.
%
%
%
%
% All of the research in this thesis occurred on the unceded territory of the Wampanoag Nation, and I acknowledge the painful history of genocide, oppression, and forced removal of indienous people from this territory.
%
%
%
% nuh
% 
% joe, liang
% jay, sebastian, mehmet
% anshul, baiqing
% zongqi, karna
% mike maikowsky
% gerry, lars
% the gedik group
% mary ann, tom, alison, and dylan
% marc, lisa, antonella, julius, and ari
% 
% rachel, abe, jeff, cliff, will, hope, kalina, shehper
% will, cliff, kalina, hope, abe, jeff, rachel
% 
% camilla
% 
% 
% workers
% mit supercloud
% land acknowledgment
%
% Anyone who's done a Ph.D. knows it's a task filled with all sorts of challenges and, at times, hardship.
% In the beginning of my Ph.D., I fought through these hardships because I cared about science, and I wanted to contribute to it if I could, even if that ambition felt impossible at times.
% At some point, though, I realized my motivation shifted---I still care about those things, but, more immediately, I care about the people around me, and the people that look out for me.
% To abandon that ambition would be, in some sense, like letting them down.
% Here I 
% 
% Now, the feeling is different.
% These days, I do it for the people around me who support what I do.
% They look out for me, 
% 
% 
% Anyone who's had the chance to do a Ph.D. knows that the pursuit does involve some challenges and, at times, some amount of hardship as well.
% In the beginning of my Ph.D., I faced these challenges because I cared about science, and I wanted to contribute to it if I could, even if that ambition felt impossible at times.
% But at some point, my motivation shifted---I still care about those things, but, more immediately, I care about the people around me, and the people that look out for me.
% Somehow it feels like abandoning my ambitions would be, in some sense, like letting them down.
% 
% 
% First, I wish to acknowledge the painful history of genocide, oppression, and forced removal of indigenous people from the unceded territory of the Wampanoag Nation, on which all of the research presented in this thesis occurred.
% All of the research in this thesis occurred on the unceded territory of the Wampanoag Nation, and I acknowledge the painful history of genocide, oppression, and forced removal of indienous people from this territory.
%
